An empirical study of fixation for strategies in the Iterated Prisoner's Dilemma

### 2017-04-29 main.tex
Add tables showing rank vs size.

@@

Figures~\ref{fig:ranks_v_size_invade},~\ref{fig:ranks_v_size_resist}
and~\ref{fig:ranks_v_size_coexist} show the median rank of each strategy against
population size. Note that these ranks are
not necessarily integers as group ties are given the average rank.

\begin{figure}[!hbtp]
    \centering
    \includegraphics[height=.9\textheight]{./img/median_rank_vs_population_size_invade.pdf}
    \caption{invade}
    \label{fig:ranks_v_size_invade}
\end{figure}

\begin{figure}[!hbtp]
    \centering
    \includegraphics[height=.9\textheight]{./img/median_rank_vs_population_size_resist.pdf}
    \caption{resist}
    \label{fig:ranks_v_size_resist}
\end{figure}

\begin{figure}[!hbtp]
    \centering
    \includegraphics[height=.9\textheight]{./img/median_rank_vs_population_size_coexist.pdf}
    \caption{Coexistance}
    \label{fig:ranks_v_size_coexist}
\end{figure}

Tables~\ref{tbl:correlation_coeficients_invade},~\ref{tbl:correlation_coeficients_resist}
and~\ref{tbl:correlation_coeficients_coexist} show the correlation coefficients
of the ranks in of strategies in differing population size. This is shown
graphically in Figure~\ref{fig:correlation_coefficients}. It is immediate to
note that how well a strategy performs in any Moran process for \(N>2\) has
little to do with the performance for \(N=2\).
% TODO This is a really good conclusion I think: it shows that when studying
% population dynamics it is important to consider larger populations.

@@

% TODO Discuss effect of size on performance of strategies. Relate to Press and
% Dyson work: "ZD strategies win matches but not tournaments".

% TODO Discuss some particular interesting cases; strategies that perform well
% in the overall tournament for v2.8.0 perhaps?

@@

% TODO Write overall conclusion. (Perhaps: weakness of short memory
% strategies?)

Further work:

\begin{itemize}
    \item Spatial structure;
    \item More than two types in the population;
    \item Modified Moran processes (Fermi selection);
    \item Mutation;
\end{itemize}
% TODO Write this section.

>>>>

% TODO Archive all data (Zenodo) and include details/reference here.

@@

Figures~\ref{fig:ranks_v_size_invade},~\ref{fig:ranks_v_size_resist}
and~\ref{fig:ranks_v_size_coexist} show the median rank of each strategy against
population size. For all starting populations \(i\in\{1, N/2, N-1\}\) the ranks
of strategies are relatively stable across the different values of \(N>2\)
however for \(N=2\) there is a distinct difference. This confirms what has been
discussed in previous sections.

Tables~\ref{tbl:ranks_v_size_invade},~\ref{tbl:ranks_v_size_resist}
and~\ref{tbl:ranks_v_size_coexist} show the same information for the strategies
that rated high for \(N=2\) and \(N=12\).
% TODO Change to N=14?

\begin{figure}[!hbtp]
    \centering
    \includegraphics[height=.9\textheight]{./img/median_rank_vs_population_size_invade.pdf}
    \caption{Ranks of all strategies according to \(x_1\) for different
    population sizes}
    \label{fig:ranks_v_size_invade}
\end{figure}

\begin{table}[!hbtp]
    \centering
    \scriptsize
    \input{./tex/change_of_rank_invade.tex}
    \caption{Ranks of some strategies according to \(x_1\) for different
    population sizes}
    \label{tbl:ranks_v_size_invade}
\end{table}

\begin{figure}[!hbtp]
    \centering
    \includegraphics[height=.9\textheight]{./img/median_rank_vs_population_size_resist.pdf}
    \caption{Ranks of all strategies according to \(x_{N-1}\) for different
    population sizes}
    \label{fig:ranks_v_size_resist}
\end{figure}

\begin{table}[!hbtp]
    \centering
    \scriptsize
    \input{./tex/change_of_rank_resist.tex}
    \caption{Ranks of some strategies according to \(x_{N-1}\) for different
    population sizes}
    \label{tbl:ranks_v_size_resist}
\end{table}

\begin{figure}[!hbtp]
    \centering
    \includegraphics[height=.9\textheight]{./img/median_rank_vs_population_size_coexist.pdf}
    \caption{Ranks of all strategies according to \(x_{N/2}\) for different
    population sizes}
    \label{fig:ranks_v_size_coexist}
\end{figure}

\begin{table}[!hbtp]
    \centering
    \scriptsize
    \input{./tex/change_of_rank_coexist.tex}
    \caption{Ranks of some strategies according to \(x_{N/2}\) for different
    population sizes}
    \label{tbl:ranks_v_size_coexist}
\end{table}

Tables~\ref{tbl:correlation_coeficients_invade},~\ref{tbl:correlation_coeficients_resist}
and~\ref{tbl:correlation_coeficients_coexist} show the correlation coefficients
of the ranks in of strategies in differing population size. This is shown
graphically in Figure~\ref{fig:correlation_coefficients}. It is immediate to
note that how well a strategy performs in any Moran process for \(N>2\) has
little to do with the performance for \(N=2\). This illustrates why the strong
performance of zero determinant strategies predicted in \cite{Press2012} does
not extend to larger populations. This was discussed theoretically in
\cite{Adami2013} however not observed empirically at the scale presented here.

@@

@@

A detailed empirical analysis of 164 strategies of the IPD within a pairwise
Moran process has been carried out. All \(\binom{164}{2}=13,366\) possible
ordered pairs of strategies have been placed in a Moran process with different
starting values allowing the each strategy to attempt to invade the other.

This is the largest such experiment carried out and has lead to many insights.

When studying evolutionary processes it is vital to consider \(N>2\) as the
special case for \(N=2\) cannot be used to extrapolate performance in bigger
populations. This was shown both observationally in
Sections~\ref{sec:strong_invaders} and~\ref{sec:strong_resistors} but also by
considering the correlation of the ranks in different population sizes in
Section~\ref{sec:population_size}.

For \(N=2\), memory one strategies perform well, in particular as predicted by
\cite{Press2012} zero determinant strategies rank highly. However, there are no
memory one strategies in the top 5 performing strategies for \(N>3\). This is
due to their lack of sophistication which allows them to recognise and adjust to
their opponent.

It is felt that these findings are important for the ongoing understanding of
population dynamics and offer evidence for some of the shortcomings of short
memory which has started to be recognised by the community~\cite{Hilbe2017}.

All source code for this work has been written in a sustainable manner: it is
open source, under version control and tested which ensures that all results can
be reproduced \cite{Prlic2012, Sandve2013, Wilson2014}. The raw data as well as
the processed data has also been properly archived.

There are various areas for further work to build on this. Firstly, an analysis
of the effect of noise would offer insights about the stability of the findings.
It would also be possible to consider three or more types of strategy in the
population and finally mutation would also offer an interesting dimension to
explore.
